### 2023-03-24 Response.txt
Response for ICALP

We would like to thank the reviewers for their work and comments.

ANSWER TO REVIEWER 1:
Regarding the movitation of periodicity: We want to point out that, since every edge having the same period is clearly a special case of every edge having *some* period, our hardness results also hold for the more general/realistic setting that the reviewer described. An even more general setting would be to allow more than one label per edge per periods (this can model different periods by taking the LCM of them as a global period). We propose this as a future research direction in the conclusion (lines 603-604). However, we believe that thoroughly understanding the fundamental case of having the same period and exactly one label per edge per period is vital before moving towards understanding more general/realistic settings.

Regarding the motivation of considering durations: We agree that from the "design-perspective" many other properties to realize might be well-motivated, such as the ones proposed by the reviewer. However, from the "verification-perspective" (cf. lines 28-31) it is well-motivated that a matrix of e.g. measured duration-values is given, and we are asked to reconstruct the temporal network displaying exactly those durations.

Regarding the motivation of studying parameterized complexity: This is a canonical and well-established way to approach NP-hard problems when exact algorithms are desired. We did not study approximation because our problem does not have a canonical formulation as an optimization problem. What would be the objective function? If we want to get the durations as close as possible to the ones prescribed in the matrix, it opens new questions such as e.g. "How many edges are we allowed to put into the graph?". Since this is a study of a new problem, we believe it is natural to initially focus on exact algorithms.

Since we identified a restricted setting (where the underlying graph is a forest) in which our problem is solvable in polynomial time, it is natural to explore how this algorithm generalizes. This polynomial-time solvable special case of trees gives a strong (theoretical) motivation to study parameters that measure "tree-likeness". From a practical standpoint, for many real-world networks it is well-motivated to assume that they are very structured, which motivates the study of structural graph parameters. Tree-likeness is a fairly common assumption, in particular for transportation networks (see e.g. reference [19]).

Note that our main parameterized hardness result (Theorem 4) rules out the possibility of obtaining FPT algorithms for many structural graph parameters, and not only for the ones that measure tree-likeness. So the question rather becomes "Which structural properties of the underlying graph allow for efficient algorithms?". Note that e.g. the arboricity can be upper-bounded by a function of the feedback vertex number, and hence we cannot expect to obtain fixed-parameter tractability for this parameter. Note that also parameters that can be small in dense graphs, such as the cliquewidth or the vertex-deletion-distance to chordal graphs, can be upper-bounded by a function of the feedback vertex number. Hence, we do have a strong (theoretical) justification for obtaining fixed-parameter tractability with the feedback edge number (although it is a rather large parameter), because we can rule out fixed-parameter tractability for most smaller parameters (as discussed e.g. in lines 363-364). We indeed leave the parameterized complexity wrt. pathwidth open, and in the conclusion of the full version we suggest studying this (see appendix lines 2198-2199).

We thank the reviewer for pointing us towards "dynamic graphs" and we will add a discussion about this to the related work section. However, this is a related but *different* mathematical model. Note that we explicitly construct a (periodic) temporal graph, that is, the mathematical object contains all information about how the graph changes (deterministically). In dynamic graphs, as described in the work "Dynamic networks: models and algorithms" by Kuhn and Oshman, the assumption is that the changes are adversarial or random (while obeying some constraints such as connectivity) and not necessarily known in advance.

ANSWER TO REVIEWER 2:
We made great effort to provide intuitive explanations of the main ideas and techniques we use, as well as some technical details that convey an intuition about the remaining steps necessary to complete the proofs (which are given in the appendix). We believe that our work is relevant to a larger audience as it initiates the research of a well-motivated new problem setting and generates several future research directions. We see much potential for positive results.

Definition 13 defines segments as paths between two vertices of interest. Indeed, these path segments should be non-trivial (i.e. they must have at least two edges), we will clarify this in teh final version. If there were two (non-trivial) segments between two vertices of interest, then there would be a cycle which contains a feedback edge. It follows that one of the segments must contain a further vertex of interest, which is not allowed by the definition.

ANSWER TO REVIEWER 3:
Each connector gadget contributes 2 vertices to the FVS, since many other vertices are connected to \hat{v}_1 and \hat{v}_2. The edges are omitted from the picture to avoid cluttering. We will add indications of the missing edges to the figure and clarify this in the figure caption.

We provide a definition of the problem MULTICOLORED CLIQUE in lines 187-188.