### 2023-12-20 SAND2024/Figures/fig-guess1.tex

% tikzpic.tex
\documentclass[crop,tikz]{standalone}% 'crop' is the default for v1.0, before it was 'preview'

\usetikzlibrary{arrows,decorations.pathmorphing,decorations.pathreplacing,backgrounds,positioning,fit,matrix}
\usetikzlibrary{shapes,calc,patterns,arrows.meta}
\tikzset{
	vert/.style={circle,inner sep=1.5,fill=white,draw,minimum size=.3cm},
	edge/.style={color=black, thick},
	diredge/.style={->,>={Stealth[width=8pt,length=8pt]},color=black, thick},
	timelabel/.style={fill=white,font=\footnotesize, text centered},
	wave/.style={decorate,decoration={coil,aspect=0}},
	dirwave/.style={->, >={Stealth[width=8pt,length=8pt]},decorate,decoration={coil,aspect=0}},
	diredge2/.style={->,>={Stealth[width=8pt,length=8pt]}}
}
\begin{document}

			\begin{tikzpicture}
		%%%S_uv
		\node[vert,label=above:$u {=} v_1$] (v1) at (1,0) {};
		\node[vert,label=above:$v_2$] (v2) at (2,0) {};
		\node[vert,label=above:$v_{p-1}$] (vp1) at (6,0) {};
		\node[vert,label=above:$v_p {=} v$] (vp) at (7,0) {};
		\draw (v1) -- (v2) -- (vp1) -- (vp);

		%%%% S_wz
		\node[vert,label=below:$w {=} z_1$] (z1) at (1,-3) {};
		\node[vert,label=below:$z_2$] (z2) at (2,-3) {};
		\node[vert,label=below:$z_{r-1}$] (zr1) at (6,-3) {};
		\node[vert,label=below:$z_r {=} z$] (zr) at (7,-3) {};
		\draw (z1) -- (z2) -- (zr1) -- (zr);

		\draw [violet] (v2) edge[diredge2] (v1);

		\draw [blue]%[transform canvas={yshift=0.5mm},thick] 
		(v1) edge[dirwave] [out=250,in=120] (z1) 
		(z1) edge[diredge2] (z2);

		\draw [red] %[transform canvas={yshift=-0.5mm}] 
		(v1) edge[dirwave] [out=300,in=90] (zr) 
		(zr) edge[diredge2] (zr1);
	\end{tikzpicture}

\end{document}